The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes

### 2019-08-02 paper2_JNLS/paper.tex
NM

With the above choices, the dimensionless KdV takes the form
\begin{align}
\label{dimlessKdV}
&u_t + C_3 \drat^{-3/2} \, u u_x + C_2 \drat^{1/2} \, u_{xxx} = 0
\qquad \text{for } x \in [-\pi,\pi] \\
&C_3 = \frac{3}{2} \pi^{3/2} \epsup \delup^{-1} \, , \quad
C_2 = \frac{\pi^3 \delup}{6 \lamfac^2} 
\end{align}
where dimensionless parameters are given by
\begin{align}
&\drat = {\depth}/{\dup} \, .
&&\mbox{\em dimensionless depth} \\
&\epsup = \etastd / \dup
&&\mbox{\em upstream amplitude-to-depth ratio} \\
&\delup = \dup / \lamup
&&\mbox{\em upstream depth-to-wavelength ratio}
\end{align}
While $\epsup$ and $\delup$ are fixed parameters, the dimensionless depth, $\drat = \dratupdn$, changes value across the ADC. Recall that the reference frame of \eqref{dimlessKdV} moves with the local wave speed via the variable $\xi = x-ct$ from \eqref{KdV}. Thus, the ADC is met at some time $T_{ADC}$, and for simplicity we set $T_{ADC} = 0$. Then the dimensionless coefficient $\drat$ is a piece-wise constant function of dimensionless time
\begin{equation}
\drat = 
\begin{cases}
1 		&\quad \mbox{for } {t}<0 \\
\dratdn = {\ddn}/{\dup} 	&\quad \mbox{for } {t}>0
\end{cases}
\end{equation}

A few comments are in order. First, we note that the original formulation of this theory utilized a slightly different normalization \cite{majda2019statistical}, with identical powers of $\drat$ but with different expressions for the other dimensionless parameters. These differences are purely cosmetic, and we have made the choices above simply to facilitate comparison against the experiments. Second, an alternate formulation of the variable-depth KdV equation exists in which it is conjectured that the product $\depth^{1/4} \eta$, rather than $\eta$, varies continuously across the ADC \cite{johnson1997modern}. Of course, that assumption implies a discontinuity in surface displacement, which, though small, is physically unrealistic. We have chosen to enforce continuity of surface displacement in the above on the basis of physical realism. Using the alternative formulation would simply modify the power of $\drat$ in the second term of \eqref{dimlessKdV} by 1/4, i.e.~$\drat^{-7/4}$ instead of $\drat^{-3/2}$. Hence, all of the results discussed below would still hold qualitatively in this alternate formulation, with only slight quantitative modifications.

@@

The variable-depth KdV \eqref{dimlessKdV}, though not Hamiltonian throughout the entire domain, admits a Hamiltonian structure on each side of the ADC. Indeed, \eqref{dimlessKdV} can be expressed as
\begin{align}
\dot{u} = \sympJ \vard{\Ham}{u}
\end{align}
where $\sympJ = \pdi{}{x}$ is the symplectic operator and $\Ham$ the Hamiltonian, which will take different forms on either side of the ADC. It is convenient to decompose the Hamiltonian into a so-called cubic and quadratic component, given respectively by
\begin{align}
\Hthree = \frac{1}{6} \int_{-\pi}^{\pi} u^3 \dx	\, , \qquad
\Htwo = \frac{1}{2} \int_{-\pi}^{\pi} u_x^2 \dx	\, .
\end{align}
Then the Hamiltonian can be expressed as
\begin{equation}
\Ham = C_2 \drat^{1/2} \, \Htwo - C_3 \drat^{-3/2} \, \Hthree
\end{equation}
where $\drat$ changers value across the ADC. More explicitly, the separate upstream and downstream Hamiltonians are given by
\begin{align}
&\Hup = C_2 \, \Htwo - C_3 \, \Hthree  \\
&\Hdn = C_2 \dratdn^{1/2} \, \Htwo - C_3 \dratdn^{-3/2} \, \Hthree
\end{align}
The above formulation differs cosmetically from previous ones \cite{abramov2003hamiltonian, majda2019statistical}, which used the opposite sign in both $\sympJ$ and $\Ham$ and found  {\em negative} inverse temperature to be appropriate for recovering the experimental observations. We have changed the signs of both $\sympJ$ and $\Ham$ so that a {\em positive} inverse temperature could be used and thus fit the theory into the most conventional statistical mechanics.

>>>>

With the above choices, the dimensionless KdV takes the form
\begin{align}
\label{dimlessKdV}
&u_t + C_3 \drat^{-3/2} \, u u_x + C_2 \drat^{1/2} \, u_{xxx} = 0
\qquad \text{for } x \in [-\pi,\pi] \\
&C_3 = \frac{3}{2} \pi^{3/2} \epsup \delup^{-1} \, , \quad
C_2 = \frac{\pi^3 \delup}{6 \lamfac^2} 
\end{align}
The constants $C_3$ and $C_2$ do {\em not} change value crossing the ADC and are given in terms of the dimensionless parameters
\begin{align}
&\epsup = \etastd / \dup
&&\mbox{\em upstream amplitude-to-depth ratio} \\
&\delup = \dup / \lamup
&&\mbox{\em upstream depth-to-wavelength ratio}
\end{align}
The reason for the subscripts $3$ and $2$ will become evident in the next section.

Meanwhile, the dimensionless depth $\drat = {\depth}/{\dup}$ {\em does} change value across the ADC since the depth $\depth$ changes. Recall that the reference frame of \eqref{dimlessKdV} moves with the local wave speed via the variable $\xi = x-ct$ from \eqref{KdV}. Thus, the ADC is met at some time $T_{ADC}$, and for simplicity we set $T_{ADC} = 0$. Therefore, we can regard $\drat$ is a piece-wise constant function of dimensionless time
\begin{equation}
\drat = 
\begin{cases}
1 		&\quad \mbox{for } {t}<0 \\
\dratdn = {\ddn}/{\dup} 	&\quad \mbox{for } {t}>0
\end{cases}
\end{equation}

A few comments are in order. First, we note that the original formulation of this theory utilized a slightly different normalization \cite{majda2019statistical}, with identical powers of $\drat$ in \eqref{dimlessKdV} but with different expressions for the other dimensionless parameters. These differences are purely cosmetic, and we have made the choices above simply to facilitate comparison against the experiments. Second, an alternate formulation of the variable-depth KdV equation exists in which it is conjectured that the product $\depth^{1/4} \eta$, rather than $\eta$, varies continuously across the ADC \cite{johnson1997modern}. Of course, that assumption implies a discontinuity in surface displacement, which, though small, is physically unrealistic. We have chosen to enforce continuity of surface displacement in the above on the basis of physical realism. Using the alternative formulation would simply modify the power of $\drat$ in the second term of \eqref{dimlessKdV} by 1/4, i.e.~$\drat^{-7/4}$ instead of $\drat^{-3/2}$. Hence, all of the results discussed below would still hold qualitatively in this alternate formulation, with only slight quantitative modifications.

@@

The variable-depth KdV \eqref{dimlessKdV}, though not Hamiltonian throughout the entire domain, admits a Hamiltonian structure on each side of the ADC. Indeed, \eqref{dimlessKdV} can be expressed as
\begin{align}
\dot{u} = \sympJ \vard{\Ham}{u}
\end{align}
where $\sympJ = \pdi{}{x}$ is the symplectic operator and $\Ham$ the Hamiltonian, which will take different forms on either side of the ADC. It is convenient to decompose the Hamiltonian into a so-called cubic and quadratic component, given respectively by
\begin{align}
\Hthree = \frac{1}{6} \int_{-\pi}^{\pi} u^3 \dx	\, , \qquad
\Htwo = \frac{1}{2} \int_{-\pi}^{\pi} u_x^2 \dx	\, .
\end{align}
Then the Hamiltonian can be expressed as
\begin{equation}
\Ham = C_2 \drat^{1/2} \, \Htwo - C_3 \drat^{-3/2} \, \Hthree
\end{equation}
where $\drat$ changers value across the ADC. More explicitly, the separate upstream and downstream Hamiltonians are given by
\begin{align}
&\Hup = C_2 \, \Htwo - C_3 \, \Hthree  \\
&\Hdn = C_2 \dratdn^{1/2} \, \Htwo - C_3 \dratdn^{-3/2} \, \Hthree
\end{align}
The above formulation differs cosmetically from some previous ones \cite{abramov2003hamiltonian, bajars2013weakly, majda2019statistical}, which used the opposite sign in both $\sympJ$ and $\Ham$. With those signs, \cite{majda2019statistical} found  {\em negative} inverse temperature to be appropriate for recovering the experimental observations. We used the signs above so that a {\em positive} inverse temperature could be used and thus fit the theory into the most conventional statistical mechanics.
% Notes for myself on the sign change: Abramov CPAM 2003 and Bajars
